---
title: " progress-report-ryan-bogdan"
author: "Ryan Bogdan"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(stringi)
```

```{r}
url <- "https://www.baseball-reference.com/awards/hof_2025.shtml"
html <- read_html(url)
tables <- html_table(html)

head(tables[[1]], 3) # The column headers are first in row 1

data <- tables[[1]]
actual_col_names <- data[1, ]
colnames(data) <- actual_col_names
data <- data[-1, ]
head(data, 3) # Now the column Headers are not in row 1

# Select only the columns that we are interested in
data <- data[, 1:5]

# The % Vote column has the % inside the values. The % should be removed
data$`%vote` <- gsub("%", "", data$`%vote`)

# Additionally, YoB is in the format "1st" "10th", etc. This will be changed to "1", "10", etc.
data$YoB <- readr::parse_number(data$YoB)

# Additionally, some names have "X-" in front of them
data$Name <- gsub("X-", "", data$Name)

# Now the data values should be good
head(data)

str(data) # Check datatypes 
data$Rk <- as.numeric(data$Rk)
data$Votes <- as.numeric(data$Votes)
data$`%vote` <- as.numeric(data$`%vote`)
str(data) # Now numeric datatypes are numericS

library(Lahman) # To get HallOfFame dataset

# Currently some players have accents in them. Remove them so the playerID from People can be in the data dataframe
data <- data %>%
  mutate(Name = stri_trans_general(Name, "Latin-ASCII"))

# Match the playerID from People dataframe to the data dataframe
People <- People %>%
  mutate(fullName = paste(nameFirst, nameLast, sep = " "))
data <- data %>%
  left_join(People %>% select(fullName, playerID), by = c("Name" = "fullName"))

# Create a new dataframe containing the new 2025 data in the HallOfFame data format
hof2025 <- data %>%
  mutate(
    yearID = 2025,
    votedBy = "BBWAA",
    ballots = round(Votes / (`%vote` / 100)),  
    needed = round(ballots * 0.75), 
    inducted = ifelse(Votes > needed, "Y", "N"),
    category = "Player",
    needed_vote = NA 
  ) %>%
  select(playerID, yearID, votedBy, ballots, needed, votes = Votes, inducted, category, needed_vote)
colnames(hof2025) <- names(HallOfFame)
HallOfFame <- rbind(HallOfFame, hof2025)

# Output a csv
readr::write_csv(HallOfFame, file="HallOfFame.csv")

# New graph including 2025 data
HallOfFame %>% 
  ggplot(aes(x = yearID, fill = inducted)) +
  geom_bar() +
  xlim(c(1936, 2025))
```

Problems encountered: After following the steps on the lab 4 html github slides, the first problem I ran into was when trying to merge the People playerID into the data dataframe. Some of the playerID's were NA. This was because some of the names had accents in them in the data dataframe, but didn't have them in the People dataframe. I was able to resolve this issue. Another problem I had was when I tried to run rbind to append my 2025 results into the existing HallOfFame dataframe. I was able to solve this issue by running "colnames(hof2025) <- names(HallOfFame)", which essentially made the column names the same.
