---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(stringi)
```

# Lab report #4 - instructions

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday (after Thanksgiving) to polish things.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 4: Scraping (into) the Hall of Fame

```{r}
url <- "https://www.baseball-reference.com/awards/hof_2025.shtml"
html <- read_html(url)
tables <- html_table(html)

head(tables[[1]], 3) # The column headers are first in row 1

data <- tables[[1]]
actual_col_names <- data[1, ]
colnames(data) <- actual_col_names
data <- data[-1, ]
head(data, 3) # Now the column Headers are not in row 1

# Select only the columns that we are interested in
data <- data[, 1:5]

# The % Vote column has the % inside the values. The % should be removed
data$`%vote` <- gsub("%", "", data$`%vote`)

# Additionally, YoB is in the format "1st" "10th", etc. This will be changed to "1", "10", etc.
data$YoB <- readr::parse_number(data$YoB)

# Additionally, some names have "X-" in front of them
data$Name <- gsub("X-", "", data$Name)

# Now the data values should be good
head(data)

str(data) # Check datatypes 
data$Rk <- as.numeric(data$Rk)
data$Votes <- as.numeric(data$Votes)
data$`%vote` <- as.numeric(data$`%vote`)
str(data) # Now numeric datatypes are numericS

library(Lahman) # To get HallOfFame dataset

# Currently some players have accents in them. Remove them so the playerID from People can be in the data dataframe
data <- data %>%
  mutate(Name = stri_trans_general(Name, "Latin-ASCII"))

# Match the playerID from People dataframe to the data dataframe
People <- People %>%
  mutate(fullName = paste(nameFirst, nameLast, sep = " "))
data <- data %>%
  left_join(People %>% select(fullName, playerID), by = c("Name" = "fullName"))

# Create a new dataframe containing the new 2025 data in the HallOfFame data format
hof2025 <- data %>%
  mutate(
    yearID = 2025,
    votedBy = "BBWAA",
    ballots = round(Votes / (`%vote` / 100)),  
    needed = round(ballots * 0.75), 
    inducted = ifelse(Votes > needed, "Y", "N"),
    category = "Player",
    needed_vote = NA 
  ) %>%
  select(playerID, yearID, votedBy, ballots, needed, votes = Votes, inducted, category, needed_vote)
colnames(hof2025) <- names(HallOfFame)
HallOfFame <- rbind(HallOfFame, hof2025)

# Output a csv
readr::write_csv(HallOfFame, file="HallOfFame.csv")

# New graph including 2025 data
HallOfFame %>% 
  ggplot(aes(x = yearID, fill = inducted)) +
  geom_bar() +
  xlim(c(1936, 2025))
```
